Demand Forecasting of Individual Probability Density Functions with Machine Learning

### 2020-12-20 LaTex/paper.tex
included temporal confounding, reviewed intro and lit rev

% Springer
%\documentclass[smallextended]{svjour3}

@@

\newcommand{\fig}{Fig.}
\newcommand{\eqn}{Eqn.}
\newcommand{\tab}{Tab.}
\newcommand{\wrt}{w.r.t.}
\newcommand{\etal}{ {\em et al.}}

@@

\title{Demand Forecasting of individual Probability Density Functions with Machine Learning}

%Springer
%\author{
%\underline{Felix Wick} \and
%Ulrich Kerzel \and
%Martin Hahn \and
%Moritz Wolf \and
%Trapti Singhal \and
%Michael Feindt}
%
%\institute{ 
%F. Wick (corresponding author) \at 
%Blue Yonder GmbH \\ \email{felix.wick@blueyonder.com}
%\and
%U. Kerzel \at 
%IUBH Internationale Hochschule \\ \email{u.kerzel@iubh-fernstudium.de}
%\and
%M. Hahn \at
%Blue Yonder GmbH \\ \email{martin.hahn@blueyonder.com}
%\and
%M. Wolf \at 
%Blue Yonder GmbH \\ \email{moritz.wolf@blueyonder.com}
%\and
%T. Singhal \at
%Blue Yonder India Private Limited \\ \email{trapti.singhal@blueyonder.com}
%\and
%M. Feindt \at
%Blue Yonder GmbH \\ \email{michael.feindt@blueyonder.com}
%}

\author[1]{Felix Wick } %\thanks{felix.wick@blueyonder.com}}
\author[3]{Ulrich Kerzel }%\thanks{u.kerzel@iubh-fernstudium.de}}
\author[1]{Martin Hahn }%\thanks{martin.hahn@blueyonder.com}}
\author[1]{Moritz Wolf }%\thanks{moritz.wolf@blueyonder.com}}
\author[2]{Trapti Singhal }%\thanks{trapti.singhal@blueyonder.com}}
\author[1]{Michael Feindt }%\thanks{michael.feindt@blueyonder.com}}
\affil[1]{\small Blue Yonder GmbH (Ohiostra{\ss}e 8, 76149 Karlsruhe, Germany)}
\affil[2]{\small Blue Yonder India Private Limited (Bengaluru, India)}
\affil[3]{\small IUBH Internationale Hochschule (Erfurt, Germany)}

\date{}
% Springer
%\titlerunning{Demand Forecasting as a Probability Distribution}
%\authorrunning{F. Wick et al.}

@@

\begin{abstract}
Demand forecasting is a central component for many aspects of supply chain operations, as it provides crucial input for subsequent decision making like ordering processes. While machine learning methods can significantly improve prediction accuracy over traditional time series forecasting, the calculated predictions are often just point estimations for the conditional mean of the underlying probability distribution, and the most powerful approaches, like deep learning, are usually opaque in terms of how its individual predictions can be interpreted. Using the novel supervised machine learning method ``Cyclic Boosting'', complete individual probability density functions can be predicted instead of single numbers. While metrics evaluating point estimates are widely used, methods for assessing the accuracy of predicted distributions are rare and this work proposes new techniques for both qualitative and quantitative evaluation methods. Additionally, each single prediction obtained with this framework is explainable. This is a major benefit in particular for practitioners, as this allows them to avoid ``black-box'' models and understand the contributing factors for each individual prediction. 
\end{abstract}

{Keywords: \textbf{explainable machine learning, demand forecasting, probability distribution}}
%Springer
%\keywords{explainable machine learning \and demand forecasting \and probability distribution \and quantile plot}

\section{Introduction}
\label{sec:intro}

Demand forecasting is one of the main challenges for retailers and is at the core of supply chain business operations. Due to its stochastic nature, demand is difficult to forecast: It depends on many influencing factors and can be interpreted as a random variable that is described by an appropriate probability density function (PDF). Demand estimation is further complicated by the fact that retailers typically only observe realized sales rather than the actual demand, and, in case the demand exceeds the current stock level, the data become censored. Demand forecasting is an application of time series forecasting and it is important to note that demand as a random variable is not identically and independently distributed (i.i.d.) between different items and sales locations. While the probability distribution describing the demand can be attributed to a given family or parameterization, the exact parameters vary: Seasonal effects, finite life cycles of products, and the introduction of new products influence the demand distribution, as well as the local weather at the retail location or the retail location itself in terms of size, assortment range, customer diversity, and other factors. The retailers themselves also actively influence demand by using advertisements to highlight products, offering rebates or discounts for specific products, as well as pursuing an active pricing strategy. This means that while we can generally assume that demand follows a specific type of probability distribution, its parameters are unique to the instance for which an estimate is required. For example, the probability distribution governing the demand of a particular item is specific to the item, date, and retail location for which the forecast is made, and depends on a wide range of further influencing factors.

Because of the access to  Big Data, a vast amount of these data associated with demand, such as historic sales records, information about promotional events or advertisements, pricing information, local weather at retail locations, seasonal information, as well as many further variables, can be captured, stored, and processed. Different methods, such as traditional univariate time series forecasting or modern supervised machine learning algorithms, can then be used to predict the demand distribution for individual products sold at specific locations, corrected for censored data. Demand forecasting for retailers often requires to predict millions of item-store combinations daily, and traditional univariate time series forecasting methods operate on each of these many individual time series separately. Machine learning, on the contrary, optimizes all individual time series together, improving the generalizability of the method by exploiting common aspects between them. This leads to a drastic reduction in variance of the model, and in turn to an improvement of the forecast quality of each of the individual time series. Furthermore, machine learning enables a natural way to include the many exogenous variables influencing demand as features of the model, what reduces also the bias compared to traditional time series forecasting approaches which mainly rely on the historic sales.

In order to make operational decisions, an optimal demand point estimator has to be defined, that can be used to derive ordering decisions in the replenishment process of the retailer. The ordering decision process is complicated by a range of factors: Even in the case of perfect demand forecasts, the decision maker has to consider lot sizes defined by the wholesaler or manufacturer as well as to balance conflicting metrics to reach an optimal decision. Ordering too few items may result in out of stock situations leading to unrealized demand and unsatisfied customers. Ordering too many items results in excess inventory, which increases transport and storage costs and, in the case of perishable goods, excessive waste, as spoiled items need to be disposed of at additional cost and potentially even environmental impact. This situation is particularly noticeable in the so called ``ultra-fresh'' category, which includes items such as bakery products, ready-meals, fresh dairy products, or certain meat products such as ground meat. These items typically have a shelf-life ranging from less than a business day to a few business days at most, with a continuous spectrum in between, depending on the exact item. In many situations, additional constraints have to be considered to reach an optimal ordering decision: Delivery cycles of items may vary depending on the type of item and the wholesaler or manufacturer from which they are procured. Retailers also operate at a given service level to guarantee that a certain level of demand can be fulfilled. The exact service level typically depends on the overall business policy of the retailer and may also depend on individual products, ranging from ``never-out-of-stock'' items to a service level exceeding, e.g., 90 percent. Therefore, demand forecasts are ideally made in the form of full individual probability density distributions, from which optimal point estimators can be derived, which in turn serve as starting point to determine the subsequent ordering decisions.

The remainder of the paper is organized as follows: We first review the relevant literature and existing work in sec. \ref{sec:LitRev}. We then describe our method to predict individual negative binomial PDFs by means of a parametric approach including two distinct machine learning models for mean of variance in sec. \ref{sec:pdfEstimation}. After that, we describe novel techniques for the qualitative and quantitative evaluation of PDF predictions in sec. \ref{sec:pdfEvaluation}. And finally, we present a demand forecasting example to show an application of our methods in sec. \ref{sec:example}.

@@

Inventory management offers a rich theory and the extensive body of research can be broadly grouped into two categories, where the inventory control problem is either based on some knowledge of the underlying demand distribution or an integrated approach that seeks to map directly from the observed data to the ordering decision. More traditional inventory control systems rely on the knowledge of the demand distribution in one form or another, see e.g. \cite{silver1998} for an overview. In $(s,S)$ type inventory control systems \cite{Scarf1958}, inventory levels are monitored at regular intervals and orders are dispatched once the inventory level reaches a minimal value $s$. In case of linear holding and shortage costs, such policies are optimal \cite{Scarf1959}, although perishable goods pose more challenges, see e.g. \cite{Nahmias1973,Nahmias1975,nahmias1978}. Additionally, service level constraints can be included in these kind of inventory control systems \cite{minner2010periodic}. Perishable goods are well described by the ``newsvendor problem'' \cite{Edgeworth}, where, in the simplest case, all stock perishes at the end of the selling period (e.g. a business day). For a detailed review of the newsvendor problem see e.g. \cite{Khouja1999537}. Assuming linear underage and overage costs $b,h >0$, the optimal quantile $q_{\mathrm{opt}} = {b}/{(b+h})$ of a known demand distribution $f(D)$ can be calculated exactly. However, it should be noted that demand is not independent between different samples for a given date. For example, a promotion applied to one product can affect the sales of related products within the assortment. This implies that the demand forecasts for multiple items or a given assortment range cannot be treated as individual newsvendor-type predictions but need to be modeled holistically.

The direct approach is often referred to as ``data-driven newsvendor'' and discussed e.g. in \cite{beutel2012safety,ban2019big,bertsimas2020predictive,oroojlooyjadid2020applying}. It aims to avoid estimating the underlying probability distribution for demand and use the available data (historic sales records and further variables) to derive the operational decisions (i.e., the order quantity) directly. An overview of a range of different approaches can also be found in \cite{huber2019data}. Although the integrated approach seems preferable at first glance, since it avoids determining the full demand distribution and results directly in the desired operational decision, the indirect approach via demand forecasts offers some substantial advantages. First, demand forecasts in form of full PDFs can be used to simulate the performance of the relevant metrics on the level of individual item, and, for example, optimize the impact on business strategy decisions on conflicting metrics such as out of stock (i.e., lost sales) and waste rate. From a practitioners perspective, separating the demand forecast from the operational decisions (i.e., calculating the order quantities for the next delivery cycle) enables longer-term planning and reduces the complexity, as it avoids coupling delivery schedules of multiple wholesalers and manufacturers with the forecast of customer demand. It also allows to share long-term demand predictions with other business units or external vendors and wholesalers to ease their planning for the production and supply chain processes upstream of the retailer. From the perspective of the industrial practice of a vendor of supply chain methods and tools, modeling the demand separately from deriving the subsequent orders has the additional benefit that multiple retail chains can benefit from any improvement in the model description, even if the specific retailers are unrelated to each other. Additionally, a purely data-driven approach going from the observed data directly to the operational decision (such as the order quantity) does not allow to analyze the data-generating process, i.e., the mechanism behind the stochastic behavior of the customer demand. This is crucial, for example, if a causal analysis is planned, such as a study of the effect of promotions, advertisements, price changes, or other demand shaping factors in either Pearl's do-calculus \cite{PearlCausality} or Rubin's potential outcomes framework \cite{rubin1974estimating}. Using an operational quantity such as the order quantity will in most cases act as an insufficient proxy for the quantity of interest (customer demand) and likely lead to unnecessary causal pathways that one may not be able to fully control for.

Therefore, the main objective in solving the inventory control problem is to determine the underlying demand distribution. The simplest approach is to use the observed sales records and forecast these via sample average approximation (SAA), see, e.g., \cite{shapiro2014} for an overview, or as a time series (see, e.g., \cite{alwan2016}), for example by means of auto-regressive integrated moving average (ARIMA) \cite{BoxJenkins}, exponential smoothing \cite{ExponentialSmoothing}, or structural time series models \cite{StructuralTS}. However, these approaches do not make use of any data apart from the sales records themselves, although we know that many variables such as price, advertisements,  weather influence, and others are highly correlated with the demand. Saghafian and Tomlin \cite{saghafian2016newsvendor} propose to include partial information about the distribution in the derivation of the operational decision, i.e., the calculation of the optimal order quantity. As described in sec. \ref{sec:intro}, modern supervised machine learning techniques can overcome many of the issues of the more traditional approaches, and therefore are the method of choice in most use cases today (see, e.g., \cite{tsforecasting,ZHANG199835,rnn,transformer}).

In order to be able to fully optimize the operational decision, it is critical to reconstruct a full demand distribution. This also implies that a simple point-estimator, as provided by the most common statistical techniques and machine learning approaches, will not suffice. Instead, we need to determine the full demand distribution from data, conditional on the relevant variables such as date, location, and item, taking all auxiliary data such as article characteristics, pricing, advertisements, retail location details, etc. into account. This can be done in several ways: Quantile regression \cite{koenker2001,wen2017} can be implemented in various frameworks and used to estimate a range of quantiles for each predicted distribution, from which an empirical probability distribution can be interpolated. Using a dedicated neural network \cite{Feindt2006190}, either the full PDF or a defined range of quantiles can be calculated directly from the data for each individual prediction without assuming an underlying model. Alternatively, one can assume a given demand model and fit the model parameters instead of reconstructing the complete distribution \cite{astonpr373, SALINAS20201181}. This approach is computationally favorable and usually more robust, as fewer parameters need to be estimated. Empirically, one can determine the best fitting distribution from data \cite{adan1995}. However, given the stochastic nature of the demand, such an empirically determined distribution is not expected to be stable and prone to sudden changes. Instead, the choice of the demand distribution should be motivated by theoretic considerations. The discrete demand is typically modeled as a negative binomial distribution (NBD), also known as Gamma-Poisson distribution \cite{Ehrenberg1959,Ehrenberg1967,Ehrenberg1972,Chatfield1973,Schmittlein_1985}. This distribution arises if the Poisson parameter $\mu$ is a random variable itself, which follows a Gamma distribution. The NBD has two parameters, $\mu$ and $ \sigma^2 > \mu$, and is over-dispersed compared to the Poisson distribution, for which $\sigma^2 = \mu$. Hence, for each ordering decision, the model parameters $\mu$ and $\sigma$ need to be determined for each item at the required granularity, typically for each sales location and ordering time, depending on all auxiliary data describing article details, retail location, and influencing factors such as pricing and advertisement information.

This work demonstrates how the explainable machine learning algorithm Cyclic Boosting \cite{Wick2019} can be used to model the demand distribution at the granularity required by the retailer. Typically, this means that the full demand PDF has to be estimated per product for each sales location and opening day, conditional on a wide range of variables such as weather, prices, promotions, etc. In contrast to using a ``black-box'' machine learning model, Cyclic Boosting allows to interpret each individual prediction in terms of influence of the different features.

Additionally, we show how the cumulative distribution function (CDF) can be used to accurately asses the forecast quality of the full predicted demand distribution, including the tails of the distribution. This allows to verify that the predicted demand distribution accurately reflects the observed data and can hence be used both to derive operational decisions such as order quantities as well as strategic business decisions by the retailer, and to gain further insights into customer behavior using for example causal modeling.

>>>>

@@

\newcommand{\fig}{Fig.~}
\newcommand{\eqn}{Eqn.~}
\newcommand{\tab}{Tab.~}
\newcommand{\wrt}{w.r.t.}
\newcommand{\etal}{ {\em et al.}}

@@

\title{Demand Forecasting of Individual Probability Density Functions with Machine Learning}

% Michael am Ende als "PI"
\author[1]{Felix Wick}
\author[3]{Ulrich Kerzel}
\author[1]{Martin Hahn}
\author[1]{Moritz Wolf}
\author[2]{Trapti Singhal}
\author[1]{Daniel Stemmer}
\author[1]{Jakob Ernst}
\author[1]{Michael Feindt}
\affil[1]{\small Blue Yonder GmbH (Ohiostra{\ss}e 8, 76149 Karlsruhe, Germany)}
\affil[2]{\small Blue Yonder India Private Limited (Bengaluru, India)}
\affil[3]{\small IUBH Internationale Hochschule (Erfurt, Germany)}

\date{}

@@

\begin{abstract}

Demand forecasting is a central component of the replenishment process for retailers, as it provides crucial input for subsequent decision making like ordering processes. In contrast to point estimates, such as the conditional mean of the underlying probability distribution or confidence intervals, forecasting complete probability density functions allows to investigate the impact on operational metrics, which are important to define the business strategy, over the full range of the expected demand. Whereas metrics evaluating point estimates are widely used, methods for assessing the accuracy of predicted distributions are rare, and this work proposes new techniques for both qualitative and quantitative evaluation methods. Using the supervised machine learning method ``Cyclic Boosting'', complete individual probability density functions can be predicted such that each prediction is fully explainable. This is of particular importance for practitioners, as it allows to avoid ``black-box'' models and understand the contributing factors for each individual prediction. Another crucial aspect in terms of both explainability and generalizability of demand forecasting methods is the limitation of the influence of temporal confounding, which is prevalent in most state of the art approaches.

\end{abstract}

{Keywords: \textbf{explainable machine learning, retail demand forecasting, probability distribution, temporal confounding}}

\section{Introduction}
\label{sec:Intro}

% specific quantile
One of the main business operations of retailers is to ensure that sufficient goods are available in the store for customers to buy. This entails two parts: In a first step, we need to estimate the future customer demand to be able to judge how many goods we will need to order from the supplier or wholesaler. In the second step, aimed at inventory control, we need to place the actual orders, which is further complicated by the fact that multiple operational constraints, such as lot-sizes or delivery schedules, need to be included in the final ordering decision. Unfortunately, from the perspective of the retailer, the two parts, estimating the future demand and optimization of the order quantity, do not factorize into two separate issues, but are inextricably linked. This is because the business strategy depends on the complete chain of demand estimation and subsequent replenishment process to control the inventory. We can understand this in the following way: In the simplest case, a retailer might want to optimize their profit, or equivalently the cost, by reducing the impact of perishable items that have not been sold at the end of the selling period, as well as the lost revenue by not being able to meet the customer demand within this period. This is known as the ``newsvendor problem'' \cite{Edgeworth}, see, e.g., \cite{Khouja1999537} for a detailed review. Assuming linear overage ($h$) and underage ($b$) cost, the optimal quantile $q_{\mathrm{opt}} = {b}/{(b+h})$ of a known probability density function (PDF) describing demand as random variable can be calculated exactly. This quantile then corresponds to the point estimator corresponding to the numerical value of the demand prediction that can be used to determine the optimal order quantity.

% simulations
In reality, the situation is often more complex, as many factors may contribute to the cost function. Furthermore, the cost function may vary between individual stock-keeping units (SKU) and may also change with time. However, the general principle remains: If we have access to the relevant cost function, we can determine the optimal point estimator of the future demand from the associated PDF. Appendix \ref{sec:CostQuantile} demonstrates how the optimal point estimator can be calculated using the full PDF. For the case of a quadratic cost function, such as the mean squared error (MSE), or the linear cost function used in the newsvendor setting, the calculations are straight-forward. If the cost function becomes more complex, we may no longer be able to perform the calculations analytically, but will have to resort to numerical approaches. Furthermore, if we can predict more than one quantile, we can investigate the effect of choosing different points, and determine, for example, how sensitive operational metrics, such as stockout or waste rate of perishable goods, change when choosing a different quantile. This can be done using, for example, a business oriented simulation that starts from the quantiles of the predicted demand for each SKU to simulate the value of the business relevant KPIs for each setting, taking operational constraints into account. This can then be used, for example, to align the business strategy. Predicting the full demand PDF therefore allows the maximum flexibility for further optimizations, because we have direct access to all quantiles.

% PDF predictions
Due to its stochastic nature, demand is difficult to forecast: It depends on many influencing factors and, as already stated above, we can interpret it as a random variable with associated PDF. And as such, demand is neither identically and independently distributed (i.i.d.) between different products and sales locations, nor between different times. This means that, while we can generally assume that demand follows a specific type of PDF, its parameters are unique to the instance for which an estimate is required. For example, the PDF describing the demand of a particular product is specific to this product, date, and retail location for which the forecast is made. Furthermore, this distribution depends on a wide range of influencing factors, such as seasonal effects, sales price, local weather, etc. Therefore, in order to predict the expected demand for each SKU at each sales location and sales day, we need to forecast individual PDF predictions rather than globally adjusted parameters for groups of SKUs or locations or sales days.

% temporal confounding
If we account for the cases where demand exceeds the current stock level, resulting in out-of-stock situations, the actual demand in the past can be approximated by realized sales. We can then understand the time-ordered observations of sales events as a time series of a random variable. It is therefore tempting to use established time series regression models to predict the future values of the expected demand. However, if we wish to use the predicted values for anything but the prediction itself, we have to be careful to avoid, or at least limit as much as possible, temporal confounding, which emerges if some external variable $X$ causally influences several time steps $t$ of the observed variables $Y$ used in the time series, say $Y_{t}$ and $Y_{t-\tau}$ with lag $\tau$. In the case of a retailer, $Y_{t}$ is the demand for a specific product at a given sales location for a specific day, and $X$ can be, for example, the price of the product which we can set as part of our business strategy or change according to some promotion. This dependency creates a backdoor path between $Y_{t-\tau}$ and $Y_{t}$, that in turn leads to a spurious autocorrelation between the values of $Y$ at different times $t$, if we cannot control for the confounder $X$. This is particularly important for the case of demand forecasting, since customer demand is not a natural phenomenon we observe as an outsider, but retailers actively influence the customer demand  through a wide range of interventions, such as a given assortment choice, promotions, advertisements, coupons, etc. If we cannot disentangle the fundamental causal dependencies from the spurious autocorrelations, understanding the influences that lead to a specific prediction becomes very challenging, even if the underlying prediction method itself is fully explainable. Moreover, due to this disguising of actual causal dependencies, reliance on temporal confounding is also detrimental in terms of generalizability of the forecasting model. The issue of temporal confounding, together with a proposed solution, is discussed in more detail in sec. \ref{sec:ts_confounding}.

% ML
Demand forecasting for retailers often requires to predict millions of product-location combinations for each SKU each day. Univariate time series forecasting methods operate on each of these many individual time series separately. On the other hand, supervised machine learning (ML) algorithms optimize all individual time series together, which can enhance the generalizability of the method by exploiting common aspects between the various time series. This can lead to a drastic reduction in variance of the model, and in turn to an improvement of the forecast quality of each of the individual time series. Furthermore, machine learning provides a straight-forward way to include the many exogenous variables influencing demand as features (or covariates) of the model. This can also help to improve the forecast quality compared to methods mainly relying on autocorrelation of the historic sales.

Hence, in order for a retailer to be able to use demand forecasts most effectively, we need to be able to:
\begin{itemize}
	\item {Predict the future demands on SKU level at required temporal granularity as complete individual PDFs.}
	\item {Avoid temporal confounding in the model used for the prediction of the future demand.}
	\item {Evaluate that the functional form of the PDF is correct.}
	\item {Evaluate that the shape of the PDF is correct and that each part of the PDF models the observed data appropriately.}
	\item {Ideally, each prediction is fully explainable, enabling operational experts to exactly understand what contributed to it.}
\end{itemize}

The remainder of the paper is organized as follows: We first review the relevant literature and existing work in sec. \ref{sec:LitRev}. We then describe our method to predict individual negative binomial PDFs using  a parametric approach including two distinct machine learning models for mean of variance in sec. \ref{sec:pdfEstimation}. After that, we describe novel techniques for the qualitative and quantitative evaluation of PDF predictions in sec. \ref{sec:pdfEvaluation}. This is followed by a detailed discussion on temporal confounding in sec. \ref{sec:ts_confounding}. And finally, we present a demand forecasting example to show an application of our methods in sec. \ref{sec:example}.

The contribution of this paper is two-fold:

Using the explainable machine learning algorithm Cyclic Boosting \cite{Wick2019}, we show how complete probability distributions of the future demand of individual products in retail stores can be predicted at the operationally required granularity. In particular, we show how using machine learning, the detrimental effects of temporal confounding, that are intrinsic to many methods used for time series forecasting, can be limited. The impact of temporal confounding in the context of time series forecasting has not been widely investigated so far in the scientific literature, and to the best of our knowledge this is the first discussion within the context of demand forecasting.

Furthermore, we show how predictions of full individual probability density functions can be evaluated qualitatively and quantitatively, such that we can assess if both overall functional form as well as specific quantiles are predicted correctly. This allows to verify that the predicted demand distribution appropriately reflects the observed data in terms of model choice and forecast accuracy.

The general methods developed in this paper can be applied to any setting that requires the prediction and evaluation of probability distributions from observational data, not just demand forecasting. However, to make the impact and benefits more tangible in a practical setting, we limit the discussion to demand forecasting in retail.

@@

Demand forecasting is one of the core operational activities of any customer-facing business and accurate demand predictions are vital to the optimization of the business operations. This is of particular concern for example for supermarkets, as their profit margin tend to be very low, e.g., 1\% and lower for European discounters \cite{MarginLidlSweden} to about 4\% for regular supermarkets \cite{MarginWillysSweden,MarginHemkopSweden} to about 8\% in the USA \cite{MarginPublixUSA}. It can be approached in a variety of ways, see, e.g., \cite{beheshti2015survey} for a survey of methods used in retail sales forecasting with a particular focus on fashion retail.

% univariate time series methods
A popular approach is to treat the demand as a time series, i.e., as a time-ordered sequence of observed sales events. In this setting, the future demand can be predicted as a point estimate using auto-regressive integrated moving average (ARIMA) \cite{BoxJenkins} or exponential smoothing \cite{croston1972forecasting, holt1957forecasting, Brown1963, Gardner1985} approaches. See also \cite{deGooijerHyndman2006} for a comprehensive overview over traditional time series forecasting methods. Essentially, these methods extrapolate the future values by building some form of regression model based on past observations, i.e., $y_{t+1} = f(y_t, y_{t-1}, \ldots)$, introducing dedicated functional components for trends, seasonality or external factors in various ways, depending on the exact modeling approach taken. The main underlying concept behind these approaches is to exploit the autocorrelation between values of the variable $y_t$ that is to be forecasted with some lag $\tau$ between subsequent events. Intuitively, we assume that  the sequence of events observed so far is a good predictor of future events. Examples for these methods in the context of demand forecasting can be found in, e.g., \cite{huber2017cluster, kalchschmidt2006forecasting, fattah2018forecasting, permatasari2018sales}.  A different approach is to use a concept from engineering, the Kalman filter \cite{kalman1960new}, and apply it to statistical forecasting \cite{morrison1977kalman}. Examples for this method used to estimate demand are, e.g., \cite{mitropoulos1980using, tegene1991kalman, kandananond2014applying,jacobi2007water}. Starting from the concept of Kalman filters, state space models \cite{hyndman2008forecasting} relate the measured values over time to the evolution of (unknown) states of the considered system. A comparison of the performance of ARIMA and state space models in a retail setting can be found, e.g., in \cite{ramos2015performance}. Structural time series models \cite{StructuralTS} take a similar approach and decompose the observations into components for trend, seasonality, and others. A modern implementation is Facebook's ``Prophet''  \cite{taylor2018forecasting}. Several other approaches are known in the literature as well: K\"ok and Fisher \cite{kok2007demand} model the demand based on the average demand for a product, modified by a logistic regression for the number of customers in a store, purchase incidence, and product choice. Wang \etal \cite{wang2005demand} model the demand under the assumption of a Poisson process.

% ML
Alternatively to the methods discussed above, time series forecasts in general and demand predictions in particular can be approached using supervised machine learning methods, using the values of the time series to be predicted as target, i.e., the numbers the machine learning algorithm is trained to predict.  For example, see \cite{Zhang2012,remus2001neural} for applications of neural networks for time series forecasting. Machine learning approaches can include multiple covariates. This allows to include not only past sales data, but also information about prices, promotions, or details about the article or store in the forecast. There are two main approaches for time series forecasting using  machine learning: We either assume that the variables are independent random variables (for both prediction target and potentially time-dependent features) for each time step of the different time series, or we consider the full sequences of time steps as a whole, resulting in a single sample (including prediction target and potentially time-dependent features) for each individual time series. For the case of individual samples for different time steps, in order to combine the endogenous information from the target autocorrelation with all other exogenous features, the most prominent approach is to include lagged target information explicitly via stacking of univariate time series predictions for the target as features in the machine learning model. For example, we could add exponentially smoothed moving averages with a range of dampening factors as features into the model. An example for the estimation of future demand in the context of price optimization, using regression trees \cite{breiman1984}, can be found in \cite{ferreira2016analytics}. For the case of sequence samples, deep learning methods such recurrent neural networks (RNN) \cite{rnn}, especially in form of ``long short-term memory'' (LSTM) networks \cite{hochreiter1997long}, or transformers \cite{transformer}, based on the self-attention mechanism, are used. These methods use information about the prediction target autocorrelation implicitly via the input sequences of the time-ordered training events. Both concepts, LSTM cells and self-attention, aim at learning from the data the elements of the individual time series that are deemed important for the prediction of the future values by the algorithm, and implicitly rely on the autocorrelation between the observed events at different times as well. The interest in deep learning approaches for demand forecasting has increased significantly recently, see, e.g., \cite{bandara2019sales,yu2017application, goyal2018solution,helmini2019sales, golkabek2020demand}. A review of deep learning methods for time series forecasting in a wide range of applications can also be found in \cite{langkvist2014review}, the use of RNNs with focus on industrial applications is discussed in \cite{dixon2020industrial}. A further approach is to explore the use of generative adversarial networks \cite{goodfellow2014generative} in the context of time series forecasting \cite{haas2020statistical, ramponi2018t, smith2020conditional}.

% temporal confounding
In many situations, relying on the exploitation of autocorrelation, either via traditional time series forecasting or machine learning approaches, is sufficient. However, this can lead to shortcomings in terms of explainability and generalizability due to temporal confounding. There are two main paths to avoid this: One way is to explicitly account for the confounders and extend the model to include them. The main challenge is then the discovery of temporal confounders and, for example, try to learn the causal structure from the observational data, see, e.g., \cite{malinsky2018causal,runge2018causal,Runge2019}. This topic is of particular importance in fields like medicine \cite{bica2020time} or climate research \cite{perrakis2014controlling}. Brodersen \etal \cite{Brodersen2015,GoogleCausalImpact} propose to use Bayesian structural time series models to estimate the causal effect of interventions on a time series by creating a counterfactual prediction that describes the behavior of the time series had the intervention not taken place. While structural time series equations allow to model components such as trend, seasonality, and others explicitly, the fundamental form of the state equation is $Y_t = G_t Y_{t-1} + \epsilon_t$, where $\epsilon_t \sim \mathcal{N}(0,\sigma^2)$ is a white noise term and $G_t$ some matrix that describes the evolution from the state at $Y_{t-1}$ to $Y_t$. Therefore, great care must be taken to avoid temporal confounding when setting up the states for structural time series models. Compared to fields like medicine, marketing, and many others that focus on the causal impact of few or just one intervention(s), retailers have to include many external effects that affect the predicted demand and are repeated frequently: Prices are changed dynamically, advertisements placed, new products are included in the assortment and others removed, etc.

\noindent
The alternative to the above approaches is to separate the autocorrelation from exogenous information via a chain of independent models (see sec. \ref{sec:ts_confounding}).

% PDF predictions
Common to all approaches discussed so far is that they predict a point estimate, essentially a single number that indicates the next value in the sequence, for example, the number of a specific product likely to be sold in a given store on a particular opening day. While operational decisions can be based on such a single number, this does not allow us to evaluate the uncertainty on such a prediction. Since the model parameters are not {\em a priori} known but estimated from data, the predicted values are generally associated with an uncertainty that needs to be quantified. Several methods have been developed, and Chatfield \cite{chatfield1993calculating} reviews the methods available at the time. In particular, bootstrapping methods for auto-regressive time series forecasting have been proposed, see, e.g.,\cite{masarotto1990bootstrap,mccullough1994bootstrapping,mccullough1996consistent,grigoletto1998bootstrap,thombs1990bootstrap,clements2001bootstrapping,Angus1994,pascual2004bootstrap,pascual2001effects,pascual2005bootstrap}. However, as already pointed out in \cite{chatfield1993calculating}, estimating the uncertainty of the point forecast is only one reason to go beyond a point estimate. For example, if we want to be able to evaluate strategies for a range of different outcomes, we need access to more information. In particular, if we do not wish to make the assumption that the point forecast is the mean of an underlying Gaussian distribution, that can be fully specified by the point forecast and an interval, we need to be able to compute any quantile of the distribution or estimate the full PDF directly.

% quantile regression
Generally, quantile regression \cite{koenker2001,wen2017} can be implemented in various frameworks and used to estimate a range of quantiles for each predicted distribution, from which an empirical probability distribution can be interpolated. Using a dedicated neural network \cite{Feindt2006190}, either the full PDF or a defined range of quantiles can be calculated directly from the data for each individual prediction without assuming an underlying model. Focusing on time series, Hyndman \cite{hyndman1995highest} proposes to use either a simulation or bootstrapping framework to estimate the highest density forecast regions of non-normal time series. Tay \cite{tay2000density} summarizes the use of density forecasts in applications in finance and economics. Using RNNs, quantile regression for multiple forecasting horizons \cite{wen2017multi} has been used to predict the demand at the internet retailer Amazon. Using an attention-based architecture, the Temporal Fusion Transformer (TFT) \cite{lim2019temporal} can also predict a set of specified quantiles at each predicted time step.

 % PDF assumption
Alternatively, we can start from a model assumption for the demand distribution and fit the model parameters instead of reconstructing the complete distribution \cite{astonpr373, SALINAS20201181}. This approach is computationally favorable and usually more robust, as fewer parameters need to be estimated. Empirically, we can determine the best fitting distribution from data \cite{adan1995}. However, given the stochastic nature of the demand, such an empirically determined distribution is not expected to be stable and prone to sudden changes. Instead, the choice of the demand distribution should be motivated by theoretic considerations. The discrete demand is typically modeled as a negative binomial distribution (NBD), also known as Gamma-Poisson distribution \cite{Ehrenberg1959, Ehrenberg1967, Ehrenberg1972, Chatfield1973, Schmittlein_1985}. This distribution arises if the Poisson parameter $\mu$ is a random variable itself, which follows a Gamma distribution. The NBD has two parameters, the mean $\mu$ and the variance $ \sigma^2 > \mu$, and is over-dispersed compared to the Poisson distribution, for which $\sigma^2 = \mu$. Hence, for each ordering decision, the model parameters $\mu$ and $\sigma^2$ need to be determined at the required granularity, typically for each item, sales location, and ordering time, depending on all auxiliary data describing article details, retail location, and influencing factors such as pricing and advertisement information.
